Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-06-27 Paper.Rmd
summary table and sample plots inclusion

@@

tab_multivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

>>>>

\begin{figure}[!h]
\textbf{ \hspace{4.3 cm} (A) \hspace{4.3 cm} (B) \hspace{4 cm} (C) } \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=5 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}\\
\end{center}
\textbf{ \hspace{4.3 cm} (D) \hspace{4.3 cm} (E) \hspace{4 cm} (F) } \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=5 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs. ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits.
Lower panel \textbf{(D - F)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for last two datasets (see \nameref{Model_selection} of manuscript). \label{fig_biv_raw_cop}}
\end{center}
\end{figure}

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5). Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript). \label{fig_aphid_raw_cop}}
\end{center}
\end{figure}

@@

tab_biivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

### 2018-06-27 SupportingInformation.Rmd
summary table and sample plots inclusion

}
#----------------------------------
#calling the above function on 3 multivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke)
m2<-bivar_tab_summary_fn(r=BivMS_BMR)
m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67)
tab_bivar_summary<-cbind(m1,m2,m3)
colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```
```

@@

  m<-matrix(NA,1,12)
  colnames(m)<-c("#pairs","#non-indep","#non-N",
                 "Avg.best.AICw","Avg.N.AICw","#good.fit",
                 "#LT>0","#UT>0","#LT-UT>0",
                 "CI_Pl","CI_Pu","CI_Pl-Pu")
  m<-as.data.frame(m)

@@

  d<-dim(r$gfc_numBS)[1]
  m$`#pairs`<-(d^2)-d
  m$`#non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
  m$`#non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
  normal_AICw[normal_AICw==Inf]<-NaN
  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
  temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
  m$`#good.fit`<-dim(temp)[1]
  m$`#LT>0`<-sum(r$LTdep_AICw>0,na.rm=T)
  m$`#UT>0`<-sum(r$UTdep_AICw>0,na.rm=T)
  m$`#LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
  m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))

@@

>>>>

}
#----------------------------------
#calling the above function on 3 multivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke)
m2<-bivar_tab_summary_fn(r=BivMS_BMR)
m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67)
tab_bivar_summary<-cbind(m1,m2,m3)
colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```

@@

  m<-matrix(NA,1,17)
  colnames(m)<-c("#pairs","#non-indep","#non-N",
                 "Avg.best.AICw","Avg.N.AICw","#good.fit",
                 "CI0.025_LT","CI0.975_LT","CI0.025_UT","CI0.975_UT",
                 "#LT-UT>0",
                 "CI0.025_Pl","CI0.975_Pl","CI0.025_Pu","CI0.975_Pu","CI0.025_Pl-Pu","CI0.975_Pl-Pu")
  m<-as.data.frame(m)

@@

  d<-dim(r$gfc_numBS)[1]
  m$`#pairs`<-(d^2)-d
  m$`#non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
  m$`#non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
  normal_AICw[normal_AICw==Inf]<-NaN
  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
  temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
  m$`#good.fit`<-dim(temp)[1]
  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$CI0.025_LT<-round(CI_LT[1],3)
  m$CI0.975_LT<-round(CI_LT[2],3)
  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$CI0.025_UT<-round(CI_UT[1],3)
  m$CI0.975_UT<-round(CI_UT[2],3)
  m$`#LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
  m$CI0.025_Pl<-round(rs$numericdf[5,3],3)
  m$CI0.975_Pl<-round(rs$numericdf[5,4],3)
  m$CI0.025_Pu<-round(rs$numericdf[6,3],3)
  m$CI0.975_Pu<-round(rs$numericdf[6,4],3)
  m$`CI0.025_Pl-Pu`<-round(rs$numericdf[10,3],3)
  m$`CI0.975_Pl-Pu`<-list(round(rs$numericdf[10,4],3))
  #m$`CI0.975_Pl-Pu`<-round(rs$numericdf[10,4],3)
  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))

@@

```{r sample_plot_aphid_count_ff,echo=F,results="hide"}
#-------------------------------------------------------------------------------
# code to plot raw data and copula plot for aphid count @location pair (4,11)
# bestfit cop : SJ

d1<-data_aphid_count$`Green spruce aphid`[[4]]
d2<-data_aphid_count$`Green spruce aphid`[[11]]
pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid abundance from location 4",ylab="Aphid abundance from location 11",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

source("./vivj_matrix.R")
m<-vivj_matrix(d_allsp = data_aphid_count,sp=10,i=4,j=11)
pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
     cex.lab=2,cex.axis=1.5,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l')
par(op)
dev.off()

# code to plot raw data and copula plot for aphid ff @location pair (2,5)
# bestfit cop : J
d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid first-flight from location 2",ylab="Aphid first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

m<-vivj_matrix(d_allsp = data_aphid_ff,sp=11,i=2,j=5)
pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
     cex.lab=2,cex.axis=1.5,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l')
par(op)
dev.off()
```
